---
title: 'Introduction to sparklyr'
output: html_document
---

We will largely follow Chapters 2 and 3 of Mastering Spark with R, [https://therinspark.com/](https://therinspark.com/).

First, load up required packages.

```{r echo = TRUE, warning = FALSE, message = FALSE}
library(sparklyr)
library(dplyr)
library(ggplot2)
library(knitr)
```


## Connecting

Spark is used to run data-processing algorithms across a cluster. Normally, you would connect to an external cluster. However, you will connect to your local machine as if it were a cluster. This will allow you to familiarise yourself with spark and sparklyr.

The code chunk below creates a spark connection to your machine as if it were a cluster, and then loads the mtcars dataset into a Spark dataframe. mtcars is a dataset that is built into R that we will use as an example.

```{r warning = FALSE}
sc = spark_connect(master = 'local')
cars = copy_to(sc, mtcars, overwrite = TRUE)
```

## Data input/output

### Write to a csv file

This will create a folder in your working directory called cars.csv. It contains a csv with the cars data in it.

```{r}
spark_write_csv(cars, 'cars.csv')
```

Note that running this more than once will result in an error because spark_write_csv will not overwrite a folder you have already created. You may need to delete the folder cars.csv in your working directory in order to re-run.

### Read from a csv file

```{r}
spark_read_csv(sc, 'cars.csv') %>%
  head() %>%
  kable()
```

## Data wrangling

Familiar commands from dplyr work as you would expect, but now they instead connect to Spark and would be run in parallel across the cluster.

### Create a new column.

```{r}
cars = mutate(cars, transmission = ifelse(am == 0, 'automatic', 'manual'))
```

### Select columns.

```{r}
select(cars, am, transmission) %>%
  head() %>%
  kable()
```

### Calculate the mean of each column.

```{r}
summarise_all(cars, mean, na.rm = TRUE) %>%
  kable()
```


## Plots
Creating a plot isn’t usually highly computationally demanding. Therefore, sparklyr does not have a full-fledged equivalent of ggplot. It is typically best to perform all data manipulations in Spark, then bring the result back to R using the collect() command. Finally, we use the regular ggplot package to make the graph.


```{r}
# Data manipulations are done first using spark
car_group = cars %>%
  group_by(cyl) %>%
  summarise(mpg = sum(mpg, na.rm = TRUE)) %>%
  # collect brings the Spark dataframe back to a regular R dataframe
  collect()

# Now use ggplot on the R dataframe car_group
ggplot(aes(as.factor(cyl), mpg), data = car_group) +
  geom_col(fill = 'SteelBlue') +
  xlab('Cylinders') +
  coord_flip()
```


## Models, in brief

We will go into more details about these models in coming lectures.

### OLS

```{r}
ols_model = ml_linear_regression(cars, mpg ~ hp + disp)
summary(ols_model)
```

### Logistic regression

The command ml_logistic_regression can be used to train a multinomial model, where the dependent variable has more than two categories. However, it does not report standard deviations of parameter estimates.

```{r}
lr_model = ml_logistic_regression(cars, am ~ hp + disp)
summary(lr_model)
```

The command ml_generalized_linear_regression can also be used to train a
logistic model with binary dependent variable, but **dependent variables with more than
two categories are not supported!** However, it does report standard deviations 
of parameter estimates.

```{r}
lr_model = ml_generalized_linear_regression(cars, am ~ hp + disp, family = 'binomial')
summary(lr_model)
```

### Multilayer perceptron

```{r}
mlp_model = ml_multilayer_perceptron_classifier(
  cars,
  am ~ hp + disp,
  layers = c(2, 8, 8, 2)
)
predictions = ml_predict(mlp_model, cars)

select(predictions, prediction, probability_0, probability_1) %>%
  head() %>%
  kable()
```

### Gradient boosted trees

Classification trees:

```{r}
gbt_model = ml_gradient_boosted_trees(cars, am ~ hp + disp, type = 'classification')
predictions = ml_predict(gbt_model, cars)

select(predictions, prediction, probability_0, probability_1) %>%
  head() %>%
  kable()
```

Regression trees:

```{r}
gbt_model = ml_gradient_boosted_trees(cars, mpg ~ hp + disp, type = 'regression')
predictions = ml_predict(gbt_model, cars)

select(predictions, prediction) %>%
  head() %>%
  kable()
```


### Other models

Apache Spark supports many other models - I have just chosen a few to look at more closely. I encourage you to explore others! See documentation here: [https://spark.apache.org/docs/latest/ml-classification-regression.html](https://spark.apache.org/docs/latest/ml-classification-regression.html)

## Disconnecting

The following code chunk disconnects from the cluster. You should always do this after your job has been run.

```{r}
spark_disconnect(sc)
```
